---
title: "Reproducible Research: Peer Assessment 1"
author: "Jose Mari"
date: "16 de agosto de 2015"
output: html_document
---


## Loading and preprocessing the data

    1. Load the data (i.e. read.csv())

    2. Process/transform the data (if necessary) into a format suitable for your analysis


```{r echo = TRUE}
data <- na.omit(read.csv("activity.csv", colClasses = c("numeric", "Date", "numeric")))
```


## What is mean total number of steps taken per day?

    1. Make a histogram of the total number of steps taken each day

```{r fig.width=10, fig.height=5, echo = TRUE}
totalStepsPerday <- aggregate(steps ~ date, data, sum)
hist(totalStepsPerday$steps,  breaks = 20, main="DISTRIBUTION OF STEPS PER DAY", , ylab = "FREQUENCY OF STEPS", xlab = "STEPS", col="purple")
```

    2. Calculate and report the mean and median total number of steps taken per day

* Mean Steps Per Day
```{r echo = TRUE}
mean(totalStepsPerday$steps)
```

* Median Steps Per Day
```{r echo = TRUE}
median(totalStepsPerday$steps, na.rm=FALSE)
```

## What is the average daily activity pattern?

    1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, 
    averaged across all days (y-axis)

```{r fig.width=10, fig.height=5, echo = TRUE}
plot(aggregate(steps ~ date, data, mean), col="red", xlab="DATA", ylab="STEPS", lwd = 3, main="Average daily activity pattern", type="l")
```


2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r echo = TRUE}
maxSteps <- max((aggregate(steps ~ interval, data, mean))$steps)
intervals <- aggregate(steps ~ interval, data, mean)
intervals[intervals$steps == maxSteps,]$interval
```


## Imputing missing values

    1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

    2. Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. 
    For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

    3. Create a new dataset that is equal to the original dataset but with the missing data filled in.


```{r echo = TRUE}
dataAll <- read.csv("activity.csv", colClasses = c("numeric", "Date", "numeric"))

# Get all the dates in the sample
uAll <- unique(dataAll$date)

# Get dates without missing values in the sample
u <- unique(data$date)

# Get date with missing values
emptyDates <- uAll[!(uAll %in% u)]

# Get mean values of steps per day
mn <- mean(aggregate(steps ~ date, data, mean)$steps)

# populate the missing step value with the mean value of steps per day mn obtained.
dataAll$steps[dataAll$date == emptyDates[1]] <- mn
dataAll$steps[dataAll$date == emptyDates[2]] <- mn
dataAll$steps[dataAll$date == emptyDates[3]] <- mn
dataAll$steps[dataAll$date == emptyDates[4]] <- mn
dataAll$steps[dataAll$date == emptyDates[5]] <- mn
dataAll$steps[dataAll$date == emptyDates[6]] <- mn
dataAll$steps[dataAll$date == emptyDates[7]] <- mn
dataAll$steps[dataAll$date == emptyDates[8]] <- mn
```


    4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps 
    taken per day. 
    Do these values differ from the estimates from the first part of the assignment? 
    
    * Yes, these values differ from the estimates from the first part of the assignment.  
    
    
    What is the impact of imputing missing data on the estimates of the total daily number of steps?
    
    * The distribution of the steps per day does not change. 
    These changes are reflected in the highest frequency of steps per day; since, 8 additional dates with means values of steps 
    are computed.
    


```{r fig.width=10, fig.height=5, , echo = TRUE}
totalStepsPerday <- aggregate(steps ~ date, dataAll, sum)

hist(totalStepsPerday$steps,  breaks = 20, main="DISTRIBUTION OF STEPS PER DAY", , ylab = "FREQUENCY OF STEPS", xlab = "STEPS", col="purple")

```

## Are there differences in activity patterns between weekdays and weekends?

    For this part the weekdays() function may be of some help here. Use the dataset with the filled-in missing values for this part.

    1. Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a 
    weekday or weekend day.


```{r echo = TRUE}
dataExp <- na.omit(read.csv("activity.csv", colClasses = c("numeric", "Date", "numeric")))
dataExp$day <- as.POSIXlt(dataExp$date,format="%Y-%m-%d")$wday

weekday <- subset(dataExp, day != 0 & day != 6)
weekday[4] <- "weekday"

weekend <- subset(dataExp, day != 1 & day != 2 & day != 3 & day != 4 & day != 5)
weekend[4] <- "weekend" 
dataExpand <- rbind(weekday, weekend)
```


    2. Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) 
    and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 
    The plot should look something like the following, which was creating using simulated data:


```{r fig.width=10, fig.height=10, echo = TRUE}
# Create a multi-paneled plotting window
par(mfrow=c(2,1))

# draw 1,1
plot(aggregate(steps ~ interval, subset(dataExpand, dataExpand$day == "weekend"),mean), col="black", xlab="Interval", ylab="Number of steps", lwd = 2, main="weekend", type="l")

# draw 2,1
plot(aggregate(steps ~ interval, subset(dataExpand, dataExpand$day == "weekday"),mean), col="black", xlab="Interval", ylab="Number of steps", lwd = 2, main="weekday", type="l")
```

